---
title: "LAB9"
author: "Jonathan Ifegunni"
date: "October 18, 2018"
output: 
  html_document:
    fig_caption: yes
    keep_md: yes
    number_sections: yes
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#TASK 1
```{r}
getwd()
```

#TASK 2
## 
###  LINE A
Line A takes n*iter samples from the random variable x with replacement.

### Line B
Calculates the confidence interval of the statistic vector, stat. These are the values that fall between alpha/2 and 1-alpha/2

### Explain what you see (replace=TRUE)

```{r}
set.seed(35) # This will give everyone the same sample
sam=round(rnorm(20,mean=10,sd=4),2)
unique(sample(sam,20,replace=TRUE) ) # repeat this line 5Xs
```

> 
  I noticed that each time I reran the last line, the number of samples generated changed from 12-14-14-15-11 samples. Also, all the samples are unique.

### Explain what you see
(replace=FALSE)
```{r}
unique(sample(sam,20,replace=FALSE) ) # repeat this line 5Xs
```

>
  For this case, the number of samples generated remain the same for every run of the last line. Also, all the samples are unique.

### What happens and why?
```{r}
#unique(sample(sam,21,replace=FALSE) ) # repeat this line 5Xs
```

>
  R studio returns an error message.  This is because we cannot take 21 samples from a population of 20 without replacement.
 
#Task 3

##
```{r}
myboot2<-function(iter=10000,x,fun="mean",alpha=0.05,cx=1.5,...){  #Notice where the ... is repeated in the code
  n=length(x)   #sample size
  
  y=sample(x,n*iter,replace=TRUE)
  rs.mat=matrix(y,nr=n,nc=iter,byrow=TRUE)
  xstat=apply(rs.mat,2,fun) # xstat is a vector and will have iter values in it 
  ci=quantile(xstat,c(alpha/2,1-alpha/2))# Nice way to form a confidence interval
  # A histogram follows
  # The object para will contain the parameters used to make the histogram
  para=hist(xstat,freq=FALSE,las=1,
            main=paste("Histogram of Bootstrap sample statistics","\n","alpha=",alpha," iter=",iter,sep=""),
            ...)
  
  #mat will be a matrix that contains the data, this is done so that I can use apply()
  mat=matrix(x,nr=length(x),nc=1,byrow=TRUE)
  
  #pte is the point estimate
  #This uses whatever fun is
  pte=apply(mat,2,fun)
  abline(v=pte,lwd=3,col="Black")# Vertical line
  segments(ci[1],0,ci[2],0,lwd=4)      #Make the segment for the ci
  text(ci[1],0,paste("(",round(ci[1],2),sep=""),col="Red",cex=cx)
  text(ci[2],0,paste(round(ci[2],2),")",sep=""),col="Red",cex=cx)
  
  # plot the point estimate 1/2 way up the density
  text(pte,max(para$density)/2,round(pte,2),cex=cx)
  
  return(list(ci=ci,fun=fun,x=x))# Some output to use if necessary
}

################### END mybooot function #####################

# set.seed(39); sam=rnorm(25,mean=25,sd=10)
# myboot2(10000,x=sam,fun="mean",alpha=0.05,xlab="mean",col="Purple",cx=1.5)

# A
set.seed(39); sam=rnorm(25,mean=25,sd=10)
myboot2(10000,x=sam,fun="mean",alpha=0.05,xlab="mean",col="Purple",cx=1.5)
# B
set.seed(30); sam=rchisq(20,df=3)
myboot2(10000,x=sam,fun="mean",alpha=0.05,xlab="mean",col="Purple",cx=1.5)
# C
set.seed(40); sam=rgamma(30,shape=2,scale=3)
myboot2(10000,x=sam,fun="mean",alpha=0.05,xlab="mean",col="Purple",cx=1.5)
# D
set.seed(10); sam=rbeta(20,shape1=3,shape2=4)
myboot2(10000,x=sam,fun="mean",alpha=0.05,xlab="mean",col="Purple",cx=1.5)
```

##
```{r}
# means 
# for A, mean  =25
set.seed(39); sam=rnorm(25,mean=25,sd=10)
mean(sam)
# for B, mean of a chi_squared is df =3
set.seed(30); sam=rchisq(20,df=3)
mean(sam)
# C
set.seed(40); sam=rgamma(30,shape=2,scale=3)
mu_gamma = 2*3 # mean = shape * scale;
mu_gamma
# D
set.seed(10); sam=rbeta(20,shape1=3,shape2=4)
mu_beta = 3/(3+4)
mu_beta
```

>
  A -the mean of the normal distribution is 23.. Therefore, the point estimate of  23.88 is very close (it is 0.112 sd's away from the population mean)
>
  B -  the mean of the chi-squared distributed population is the degrees of freedom df =3. Therefore, the point estimate of 2.94 is very close.
>
  C - the mean of the Gamma population is 6. The point estimate is 6.5 which is still pretty close .
>
  D - the mean of the beta population is about 0.428 . The point estimate of 0.41 is a very good approximation

##
Yes, for all the generated samples, the interval contains the population values

##

```{r}

# 80% ci's
# A
set.seed(39); sam=rnorm(25,mean=25,sd=10)
myboot2(10000,x=sam,fun="mean",alpha=0.2,xlab="mean",col="Purple",cx=1.5)
# B
set.seed(30); sam=rchisq(20,df=3)
myboot2(10000,x=sam,fun="mean",alpha=0.2,xlab="mean",col="Purple",cx=1.5)
# C
set.seed(40); sam=rgamma(30,shape=2,scale=3)
myboot2(10000,x=sam,fun="mean",alpha=0.2,xlab="mean",col="Purple",cx=1.5)
# D
set.seed(10); sam=rbeta(20,shape1=3,shape2=4)
myboot2(10000,x=sam,fun="mean",alpha=0.2,xlab="mean",col="Purple",cx=1.5)
```

#TASK 4
##A)
```{r}
myboot3<-function(iter=10000,x,fun="mean",alpha=0.05,cx=1.5,...){  #Notice where the ... is repeated in the code
  n=length(x)   #sample size
  
  y=sample(x,n*iter,replace=TRUE)
  rs.mat=matrix(y,nr=n,nc=iter,byrow=TRUE)
  xstat=apply(rs.mat,2,fun) # xstat is a vector and will have iter values in it 
  ci=quantile(xstat,c(alpha/2,1-alpha/2))# Nice way to form a confidence interval
  # A histogram follows
  # The object para will contain the parameters used to make the histogram
  para=hist(xstat,freq=FALSE,las=1,
            main=paste("Histogram of Bootstrap sample statistics","\n","alpha=",alpha," iter=",iter,sep=""),
            ...)
  
  #mat will be a matrix that contains the data, this is done so that I can use apply()
  mat=matrix(x,nr=length(x),nc=1,byrow=TRUE)
  
  #pte is the point estimate
  #This uses whatever fun is
  pte=apply(mat,2,fun)
  abline(v=pte,lwd=3,col="Black")# Vertical line
  segments(ci[1],0,ci[2],0,lwd=4)      #Make the segment for the ci
  text(ci[1],0,paste("(",round(ci[1],2),sep=""),col="Red",cex=cx)
  text(ci[2],0,paste(round(ci[2],2),")",sep=""),col="Red",cex=cx)
  
  # plot the point estimate 1/2 way up the density
  text(pte,max(para$density)/2,round(pte,2),cex=cx)
  
  return(list(ci=ci,fun=fun,x=x,xstat=xstat))# Some output to use if necessary
}

sam=c(1,1,1,2,2,2,2,3,3,3,4,4)
myb3 = myboot3(10000,x=sam,fun="median",alpha=0.2,xlab="mean",col="Purple",cx=1.5)
tab = table(myb3$xstat)
barplot(tab, beside = TRUE, col = rainbow(7), main = "Barplot of xstat", xlab = "median",ylab = "Frequency")

```

##B) 
The Bootstrap interval estimate  for the median is (2,3).

#TASK 5

##
```{r}
# 95% estimate, alpha = 0.05 for fun =mean/median
myboot2<-function(iter=10000,x,fun="mean",alpha=0.05,cx=1.5,...){  #Notice where the ... is repeated in the code
  n=length(x)   #sample size
  
  y=sample(x,n*iter,replace=TRUE)
  rs.mat=matrix(y,nr=n,nc=iter,byrow=TRUE)
  xstat=apply(rs.mat,2,fun) # xstat is a vector and will have iter values in it 
  ci=quantile(xstat,c(alpha/2,1-alpha/2))# Nice way to form a confidence interval
  # A histogram follows
  # The object para will contain the parameters used to make the histogram
  para=hist(xstat,freq=FALSE,las=1,
            main=paste("Histogram of Bootstrap sample statistics","\n","alpha=",alpha," iter=",iter,sep=""),
            ...)
  
  #mat will be a matrix that contains the data, this is done so that I can use apply()
  mat=matrix(x,nr=length(x),nc=1,byrow=TRUE)
  
  #pte is the point estimate
  #This uses whatever fun is
  pte=apply(mat,2,fun)
  abline(v=pte,lwd=3,col="Black")# Vertical line
  segments(ci[1],0,ci[2],0,lwd=4)      #Make the segment for the ci
  text(ci[1],0,paste("(",round(ci[1],2),sep=""),col="Red",cex=cx)
  text(ci[2],0,paste(round(ci[2],2),")",sep=""),col="Red",cex=cx)
  
  # plot the point estimate 1/2 way up the density
  text(pte,max(para$density)/2,round(pte,2),cex=cx)
  
  return(list(ci=ci,fun=fun,x=x))# Some output to use if necessary
}
# A
set.seed(39); sam=rnorm(25,mean=25,sd=10)
myboot2(10000,x=sam,fun=function(x) mean(x)/median(x),alpha=0.05,xlab="mean(x)/median(x)",col="Purple",cx=1.5)
# B
set.seed(30); sam=rchisq(20,df=3)
myboot2(10000,x=sam,fun=function(x) mean(x)/median(x),alpha=0.05,xlab="mean(x)/median(x)",col="Purple",cx=1.5)

# C 
set.seed(40); sam=rgamma(30,shape=2,scale=3)
myboot2(10000,x=sam,fun=function(x) mean(x)/median(x),alpha=0.05,xlab="mean(x)/median(x)",col="Purple",cx=1.5)

# D 
set.seed(10); sam=rbeta(20,shape1=3,shape2=4)
myboot2(10000,x=sam,fun=function(x) mean(x)/median(x),alpha=0.05,xlab="mean(x)/median(x)",col="Purple",cx=1.5)

```

## 70% INTERVAL
```{r}
# For 70% intervals
# A
set.seed(39); sam=rnorm(25,mean=25,sd=10)
myboot2(10000,x=sam,fun=function(x) mean(x)/median(x),alpha=0.3,xlab="mean(x)/median(x)",col="Purple",cx=1.5)
# B
set.seed(30); sam=rchisq(20,df=3)
myboot2(10000,x=sam,fun=function(x) mean(x)/median(x),alpha=0.3,xlab="mean(x)/median(x)",col="Purple",cx=1.5)

# C 
set.seed(40); sam=rgamma(30,shape=2,scale=3)
myboot2(10000,x=sam,fun=function(x) mean(x)/median(x),alpha=0.3,xlab="mean(x)/median(x)",col="Purple",cx=1.5)

# D 
set.seed(10); sam=rbeta(20,shape1=3,shape2=4)
myboot2(10000,x=sam,fun=function(x) mean(x)/median(x),alpha=0.3,xlab="mean(x)/median(x)",col="Purple",cx=1.5)
```

#TASK 6
##
```{r}
? distributions
```

##
```{r}
# Uniform, Weibull, Exponential
# Generate samples of size 20
# Uniform
set.seed(39); sam=runif(20,min = 0,max = 3)
unique(sample(sam,20,replace=FALSE) )

# Weibull
set.seed(30); sam=rweibull(20,shape = 2, scale = 3)
unique(sample(sam,20,replace=FALSE) )
# Exponential
set.seed(40);sam=rexp(20,rate = 2)
unique(sample(sam,20,replace=FALSE) )
# Binomial
set.seed(10); sam=rbinom(20,40,0.6)
sam
```

##
```{r}
# 80% ci
# Uniform
sam=runif(20,min = 0,max = 3)
myboot2(10000,x=sam,fun=function(x) mean(x),alpha=0.2,xlab="mean(x)",col="Purple",cx=1.5)
myboot2(10000,x=sam,fun=function(x) var(x),alpha=0.2,xlab="var(x)",col="Green",cx=1.5)
# Weibull
sam=rweibull(20,shape = 2, scale = 3)
myboot2(10000,x=sam,fun=function(x) mean(x),alpha=0.2,xlab="mean(x)",col="Purple",cx=1.5)
myboot2(10000,x=sam,fun=function(x) var(x),alpha=0.2,xlab="var(x)",col="Green",cx=1.5)
# Exponential
sam=rexp(20,rate = 2)
myboot2(10000,x=sam,fun=function(x) mean(x),alpha=0.2,xlab="mean(x)",col="Purple",cx=1.5)
myboot2(10000,x=sam,fun=function(x) var(x),alpha=0.2,xlab="var(x)",col="Green",cx=1.5)
# Binomial
sam=rbinom(20,40,0.6)
myboot2(10000,x=sam,fun=function(x) mean(x),alpha=0.2,xlab="mean(x)",col="Purple",cx=1.5)
myboot2(10000,x=sam,fun=function(x) var(x),alpha=0.2,xlab="var(x)",col="Green",cx=1.5)
```

#TASK 7
```{r}
# first stat mean(x)/sd(x)
set.seed(68); sam=rnorm(20,mean=10,sd=4)
myboot2(10000,x=sam,fun=function(x) mean(x)/sd(x),alpha=0.2,xlab="(mean(x))/sd(x)",col="Purple",cx=1.5)
# 2nd stat IQR(x)
myboot2(10000,x=sam,fun=function(x) IQR(x),alpha=0.2,xlab="IQR(x)",col="Purple",cx=1.5)

set.seed(68); sam=rnorm(20,mean=10,sd=4)

xstat=sam# xstat is a vector and will have iter values in it 
alpha = 0.05;
ci=quantile(xstat,c(alpha/2,1-alpha/2))


set.seed(68); sam=rnorm(20,mean=10,sd=4)
mu_sam = mean(sam)
mu_sam
sd_sam = sd(sam)
sd_sam
qnorm(1-0.05/2,mean=mu_sam,sd=sd_sam)
qnorm(0.05/2, mean =mu_sam, sd =sd_sam)
myboot2(10000,x=sam,fun=function(x) mean(x),alpha=0.05,xlab="mean(x)",col="Purple",cx=1.5)
```










