---
title: "LAB8"
author: "Jonathan Ifegunni"
date: "October 11, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#TASK 1
```{r}
getwd()
```

#TASK 2
##a)
```{r}
unif = runif(10,0,5)
unif
```

##b)
```{r}
# mean and variance
mean = (0+5)/2
mean
# Variance
var = (5-0)^2/12
var
```

##c) 
```{r}
# mean/ variance of sample
mus = mean(unif)
mus
vars = (sd(unif))^2 #variance
vars
```

>
  The mean of the sample is about 1.10 more than the mean of the population, while the variance of the sample is about 0.15 more than that of the population
  
##d)
###i)
$$
E(T) = nE(Y_i)
$$
```{r}
# For the sum T
sT_mean = 10*mean 
sT_mean
```

$$
V(T) = nV(Y_i)
$$
```{r}
sT_var = 10*var
sT_var
```

###ii)

$$
E(\bar Y) = E(Y_i)
$$
```{r}
# for the mean Y
muY_mean = mean
muY_mean
```

$$
V(\bar Y) = \frac{V(Y_i)}{n}
$$
```{r}
# for the variance Y
varY_mean = var/10
varY_mean
```

##e)

###i)
This line creates n*iter uniformly distributed random variables from a = 0 to b = 5

###ii)
First, creates a matrix data specified by nr rows and nc columns and then fills the matrix by row with the values of the vector y.

###iii)
Sums up the matrix created in B across each column. Returns the result of the sums into the variable sm.

###iv)
This line is a call to the myclt function specifying values for n = 10 and iter (number of iterations) = 10000. This will run the whole function for this values.

##f)
```{r}
myclt=function(n,iter,a=0,b=5){
y=runif(n*iter,a,b)
data=matrix(y,nr=n,nc=iter,byrow=TRUE)
sm=apply(data,2,sum)
h=hist(sm,plot=FALSE)
hist(sm,col=rainbow(length(h$mids)),freq=FALSE,main="Distribution of the sum of uniforms")
curve(dnorm(x,mean=n*(a+b)/2,sd=sqrt(n*(b-a)^2/12)),add=TRUE,lwd=2,col="Blue")
sm
}
w=myclt(n=50,iter=10000,a=5,b=10)
```

##g)
```{r}
w=myclt(n=10,iter=10000) 
# sample estimates mean
mu_w = mean(w)
mu_w  
#variance
var_w = var(w)
var_w
```

##h)
```{r}
mycltm=function(n,iter,a=0,b=5){
y=runif(n*iter,a,b)
data=matrix(y,nr=n,nc=iter,byrow=TRUE)
mux=apply(data,2,mean)
h=hist(mux,plot=FALSE)
hist(mux,col=rainbow(length(h$mids)),freq=FALSE,main="Distribution of the sum of uniforms")
lines(density(mux),col="Blue",lwd=3) # add a density plot
  curve(dnorm(x,mean=(a+b)/2,sd=(b-a)/(sqrt(12*n))),add=TRUE,col="Red",lty=2,lwd=3) # add a theoretical curve
  curve(dunif(x,a,b),add=TRUE,lwd=4)
mux
}
w=mycltm(n=50,iter=10000,a=5,b=10)
```

##i)
```{r}
# sample estimates
mu_w = mean(w)
mu_w  
var_w = var(w)
var_w
```

#TASK 3
##a)
```{r}
### CLT uniform 
## my Central Limit Function
## Notice that I have assigned default values which can be changed when the function is called
mycltu=function(n,iter,a=0,b=10){
  ## r-random sample from the uniform
  y=runif(n*iter,a,b)
  ## Place these numbers into a matrix
  ## The columns will correspond to the iteration and the rows will equal the sample size n
  data=matrix(y,nr=n,nc=iter,byrow=TRUE)
  ## apply the function mean to the columns (2) of the matrix
  ## these are placed in a vector w
  w=apply(data,2,mean)
  ## We will make a histogram of the values in w
  ## How high should we make y axis?
  ## All the values used to make a histogram are placed in param (nothing is plotted yet)
  param=hist(w,plot=FALSE)
  ## Since the histogram will be a density plot we will find the max density
  
  ymax=max(param$density)
  ## To be on the safe side we will add 10% more to this
  ymax=1.1*ymax
  ## Now we can make the histogram
  hist(w,freq=FALSE,  ylim=c(0,ymax), main=paste("Histogram of sample mean",
                                                 "\n", "sample size= ",n,sep=""),xlab="Sample mean")
  ## add a density curve made from the sample distribution
  lines(density(w),col="Blue",lwd=3) # add a density plot
  ## Add a theoretical normal curve 
  curve(dnorm(x,mean=(a+b)/2,sd=(b-a)/(sqrt(12*n))),add=TRUE,col="Red",lty=2,lwd=3) # add a theoretical curve
  ## Add the density from which the samples were taken
  curve(dunif(x,a,b),add=TRUE,lwd=4)
}
```

##b)
###i)
In this case, 2 specifies that the columns should be selected.
###ii)
100,000 terms.
###iii)
plots the curve for the theoretical uniform distribution the known mean.
###iv)
This line provides the known theoretical standard deviation of the distribution,and specifies that this plot be added to a previous plot. The curve is to be plotted in red with a line width of 3.
###v)
Explain why sd takes the formula as shown in the function.
This is because the variance of the sample mean distribution is equal to variance of the original uniform distribution divided by n.

$$
Orginal\ distribution\ \sigma^2 = \frac{(b-a)^2}{12}\\
sample\ mean\ variances\ \sigma^2_m = \frac{(b-a)^2}{12n}\\
sample\ mean\ sd\ \sigma=\frac{b-a}{\sqrt{12n}}
$$

##c)
```{r}
# n=1, iter=10000, a=0,b=10
mycltu(n = 1, iter =1000, a = 0, b = 10)
# n=2, iter=10000, a=0,b=10
mycltu(n =2, iter =1000, a = 0, b = 10)
#	n=3,iter=10000,a=0,b=10
mycltu(n = 3, iter =1000, a = 0, b = 10)
#	n=5,iter=10000,a=0,b=10
mycltu(n = 5, iter =1000, a = 0, b = 10)
#	n=10, iter=10000,a=0,b=10
mycltu(n =10, iter =1000, a = 0, b = 10)
#n=30,iter=10000,a=0,b=10
mycltu(n =30, iter =1000, a = 0, b = 10)
```

##d)
I can see from the plots that the sample mean distribution obeys the central limit theorem. For a sample size of 1, the sample mean has a uniform distribution which is to be expected, since the population being sampled also has a uniform distribution. However as the sample size increases, the sample mean distribution starts to approach the normal distribution. This is clear for n =10 where the theoretical line for the normal distribution (in red) matches very closely with the density curve (blue) of the sampled distribution. Thus, the sampling distribution approaches the Normal, as n approaches infinity.

#TASK 4
```{r}
mycltb=function(n,iter,p=0.5,...){
  
  ## r-random sample from the Binomial
  y=rbinom(n*iter,size=n,prob=p)
  ## Place these numbers into a matrix
  ## The columns will correspond to the iteration and the rows will equal the sample size n
  data=matrix(y,nr=n,nc=iter,byrow=TRUE)
  ## apply the function mean to the columns (2) of the matrix
  ## these are placed in a vector w
  w=apply(data,2,mean)
  ## We will make a histogram of the values in w
  ## How high should we make y axis?
  ## All the values used to make a histogram are placed in param (nothing is plotted yet)
  param=hist(w,plot=FALSE)
  ## Since the histogram will be a density plot we will find the max density
  
  ymax=max(param$density)
  ## To be on the safe side we will add 10% more to this
  ymax=1.1*ymax
  
  ## Now we can make the histogram
  ## freq=FALSE means take a density
  hist(w,freq=FALSE,  ylim=c(0,ymax),
       main=paste("Histogram of sample mean","\n", "sample size= ",n,sep=""),
       xlab="Sample mean",...)
  ## add a density curve made from the sample distribution
  #lines(density(w),col="Blue",lwd=3) # add a density plot
  ## Add a theoretical normal curve 
  curve(dnorm(x,mean=n*p,sd=sqrt(p*(1-p))),add=TRUE,col="Red",lty=2,lwd=3) 
  
}

#	n=4,iter=10000,p=0.3
mycltb(n=4,iter=10000,p=0.3)
#	n=5,iter=10000,p=0.3
mycltb(n=5,iter=10000,p=0.3)
#	n=10,iter=10000,p=0.3
mycltb(n=10,iter=10000,p=0.3)
#	n=20,iter=10000,p=0.3
mycltb(n=20,iter=10000,p=0.3)
```

##b)
```{r}
## using p=0.7
#	n=4,iter=10000,p=0.7
mycltb(n=4,iter=10000,p=0.7)
#	n=5,iter=10000,p=0.7
mycltb(n=5,iter=10000,p=0.7)
#	n=10,iter=10000,p=0.7
mycltb(n=10,iter=10000,p=0.7)
#	n=20,iter=10000,p=0.7
mycltb(n=20,iter=10000,p=0.7)
```

##c)
```{r}
## using p=0.5
#	n=4,iter=10000,p=0.5
mycltb(n=4,iter=10000,p=0.5)
#	n=5,iter=10000,p=0.7
mycltb(n=5,iter=10000,p=0.5)
#	n=10,iter=10000,p=0.7
mycltb(n=10,iter=10000,p=0.5)
#	n=20,iter=10000,p=0.7
mycltb(n=20,iter=10000,p=0.5)
```

##d)
I can see from the plots that for the different cases of p, the sampled mean distribution approaches the normal distribution as the number of samples n gets larger. The mean of the sampled distribution increases as p or n increases. For example, for n = 5 and p = 0.5, the known theoretical mean is 2.5 and the center of the sampling distribution is at 2.5. Therefore, the sample mean distribution of a binomial population approaches the normal distribution, with a mean of np as n approaches infinity.

#TASK 5
```{r}
mycltp=function(n,iter,lambda=10,...)
  {
  ## r-random sample from the Poisson
  y=rpois(n*iter,lambda=lambda)
  ## Place these numbers into a matrix
  ## The columns will correspond to the iteration and the rows will equal the sample size n
  data=matrix(y,nr=n,nc=iter,byrow=TRUE)
  ## apply the function mean to the columns (2) of the matrix
  ## these are placed in a vector w
  w=apply(data,2,mean)
  ## We will make a histogram of the values in w
  ## How high should we make y axis?
  ## All the values used to make a histogram are placed in param (nothing is plotted yet)
  param=hist(w,plot=FALSE)
  ## Since the histogram will be a density plot we will find the max density
  
  ymax=max(param$density)
  ## To be on the safe side we will add 10% more to this
  ymax=1.1*ymax
  
  ## Make a suitable layout for graphing
  layout(matrix(c(1,1,2,3),nr=2,nc=2, byrow=TRUE))
  
  ## Now we can make the histogram
  hist(w,freq=FALSE,  ylim=c(0,ymax), col=rainbow(max(w)),
       main=paste("Histogram of sample mean","\n", "sample size= ",n," iter=",iter," lambda=",lambda,sep=""),
       xlab="Sample mean",...)
  ## add a density curve made from the sample distribution
  #lines(density(w),col="Blue",lwd=3) # add a density plot
  ## Add a theoretical normal curve 
  curve(dnorm(x,mean=lambda,sd=sqrt(lambda/n)),add=TRUE,col="Red",lty=2,lwd=3) # add a theoretical curve
  
  # Now make a new plot
  # Since y is discrete we should use a barplot
  barplot(table(y)/(n*iter),col=rainbow(max(y)), main="Barplot of sampled y", ylab ="Rel. Freq",xlab="y" )
  x=0:max(y)
  plot(x,dpois(x,lambda=lambda),type="h",lwd=5,col=rainbow(max(y)),
       main="Probability function for Poisson", ylab="Probability",xlab="y")
}

# Sampling a Poisson distribution
#	 n=2, iter=10000,lambda=4
mycltp(n=2, iter =10000,lambda =4)
#n=3,iter=10000,lambda=4
mycltp(n=3, iter =10000,lambda =4)
#n=5,iter=10000,lambda=4
mycltp(n=5, iter =10000,lambda =4)
#n=10,iter=10000,lambda=4
mycltp(n=10, iter =10000,lambda =4)
#n=20,iter=10000,lambda=4
mycltp(n=20, iter =10000,lambda =4)

# for lambda =10
#	 n=2, iter=10000,lambda=10
mycltp(n=2, iter =10000,lambda =10)
#n=3,iter=10000, lambda=10
mycltp(n=3, iter =10000,lambda =10)
#n=5,iter=10000,lambda =10
mycltp(n=5, iter =10000,lambda =10)
#n=10,iter=10000,lambda=10
mycltp(n=10, iter =10000,lambda =10)
#n=20,iter=10000,lambda=10
mycltp(n=20, iter =10000,lambda =10)


mycltp(n=10,iter=10000)
```

#TASK 6: Extra for experts
```{r}
myclts=function(n,iter,lambda=10,...){
  
  ## r-random sample from the Poisson
  y=rpois(n*iter,lambda=lambda)
  data=matrix(y,nr=n,nc=iter,byrow=TRUE)
  w=apply(data,2,sum)
  param=hist(w,plot=FALSE)
  ymax=max(param$density)
  ymax=1.1*ymax
  
  ## Make a suitable layout for graphing
  layout(matrix(c(1,1,2,3),nr=2,nc=2, byrow=TRUE))
  
  ## Now we can make the histogram
  hist(w,freq=FALSE,  ylim=c(0,ymax), col=rainbow(max(w)),
       main=paste("Histogram of sample sum","\n", "sample size= ",n," iter=",iter," lambda=",lambda,sep=""),
       xlab="Sample sum",...)
  curve(dnorm(x,mean=n*lambda,sd=sqrt(n*lambda)),add=TRUE,col="Red",lty=2,lwd=3) # add a theoretical curve
  
  barplot(table(y)/(n*iter),col=rainbow(max(y)), main="Barplot of sampled y", ylab ="Rel. Freq",xlab="y" )
  x=0:max(y)
  plot(x,dpois(x,lambda=lambda),type="h",lwd=5,col=rainbow(max(y)),
       main="Probability function for Poisson", ylab="Probability",xlab="y")
}

# Sampling a Poisson distribution
#	 n=2, iter=10000,lambda=4
myclts(n=2, iter =10000,lambda =4)
#???	n=3,iter=10000,lambda=4
myclts(n=3, iter =10000,lambda =4)
#???	n=5,iter=10000,lambda=4
myclts(n=5, iter =10000,lambda =4)
#???	n=10,iter=10000,lambda=4
myclts(n=10, iter =10000,lambda =4)
#???	n=20,iter=10000,lambda=4
myclts(n=20, iter =10000,lambda =4)

# for lambda =10
#	 n=2, iter=10000,lambda=10
myclts(n=2, iter =10000,lambda =10)
#???	n=3,iter=10000, lambda=10
myclts(n=3, iter =10000,lambda =10)
#???	n=5,iter=10000,lambda =10
myclts(n=5, iter =10000,lambda =10)
#???	n=10,iter=10000,lambda=10
myclts(n=10, iter =10000,lambda =10)
#???	n=20,iter=10000,lambda=10
myclts(n=20, iter =10000,lambda =10)
```
















