---
title: "LAB 1"
author: "Jonathan Ifegunni"
date: "August 26, 2018"
output: 
  html_document: 
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Tasks

## Task 1
```{r working directory}
getwd()
```

## Task 2
### A (Reading Data into R using the function Read.Table)
```{r Reading Data into R using the function Read.Table}
ddt=read.table(file.choose(),sep=",",header=TRUE) 
head(ddt)
```
### B (Reading Data into R using using the function Read.csv)
```{r Reading Data into R using using the function Read.csv}
ddt=read.csv("DDT.csv")
head(ddt)
```
## Task 3
### A
```{r Qualitative Varibles}
head(ddt, 3)
```
> 
     After looking at the function head() above, the following are qualitative variables
>     
     1. RIVER
     2. SPECIES


### B
```{r Quantitative variable}
head(ddt, 3)
```
> 
     After looking at the function head() above, the following are quantitative variables
>     
     1. MILE
     2. LENGHT
     3. WEIGHT
     4. DDT

### C
```{r}
summary(ddt$SPECIES)
```
### D
```{r}
with(ddt, ddt[WEIGHT>800 & SPECIES=="LMBASS",])
```
### E
```{r}
with(ddt, ddt[DDT>4.0 & RIVER=="SCM",])
```
## Clicker Questions
### Q.1
>
    The Mean Lenght of fish in the ddt dataframe is 42.8125
    
```{r}
mean(ddt$LENGTH)
```
### Q.2
>
The standard deviations of the Weight of fish is 376.5461

```{r}
sd(ddt$WEIGHT)
```


### Q.3
>
    NO - The plot shown is a plot of Weight vs Lenght and not Lenght vs Weight in the question.
    
```{r}
plot(WEIGHT~LENGTH,data=ddt)
```
    
    
### Q.4
>
   The last value of v/20 is 20
   
```{r}
v=1:20
v
```

## Task 4
### A and B
```{r}
sn=with(ddt,table(RIVER))
sn
barplot(sn, col=1:4)
```

### C and D
```{r}
spriv=with(ddt,table(RIVER, SPECIES))
spriv
barplot(spriv)
```

## TASK 5
### A
```{r}
sp=with(ddt,table(SPECIES))
pie(sp,col=1:4)
```

### B
```{r}
pie(sn,col=1:4)
```

## TASK 6

### A
```{r}
boxplot(ddt$DDT)

```

### B
```{r}
boxplot(ddt$WEIGHT)

```

### C
```{r}
boxplot(ddt$LENGTH)

```

## TASK 7

### A
```{r}
coplot(LENGTH~WEIGHT|RIVER,data=ddt)

```

### B
```{r}
coplot(DDT~WEIGHT|SPECIES,data=ddt)

```

## TASK 8
```{r}
library(ggplot2)
g= ggplot(ddt, aes(x=SPECIES, y= WEIGHT, fill = RIVER)) + geom_boxplot()+ggtitle("JONATHAN IFEGUNNI")
print(g)

```

