---
title: "Lab 5"
author: "Jonathan Ifegunni"
date: "September 18, 2018"
output:
  html_document:
    fig_caption: yes
    keep_md: yes
    number_sections: yes
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Task 1
```{r}
getwd()
```

#Task 2

##A

```{r}
## sampling function
# iter = iterations, n=sample size
# set default values
mybin=function(iter=100,n=10, p=0.5){ 
  # make a matrix to hold the samples
  #initially filled with NA's
  sam.mat=matrix(NA,nr=n,nc=iter, byrow=TRUE)
  #Make a vector to hold the number of successes in each trial
  succ=c()
  for( i in 1:iter){
    #Fill each column with a new sample
    sam.mat[,i]=sample(c(1,0),n,replace=TRUE, prob=c(p,1-p))
    #Calculate a statistic from the sample (this case it is the sum)
    succ[i]=sum(sam.mat[,i])
  }
  #Make a table of successes
  succ.tab=table(factor(succ,levels=0:n))
  #Make a barplot of the proportions
  barplot(succ.tab/(iter), col=rainbow(n+1), main="Binomial simulation", xlab="Number of successes")
  succ.tab/iter
}

```

### Ai)
>
 100 iteration
 
```{r}
windows()
mybin(iter=100,n=10, p=0.7)
```

### Aii)
>
  200 iteration

```{r}
# 200 iterations
mybin(iter=200,n=10, p=0.7)
```

### Aiii)
>
  500 iteration
  
```{r}
# 500 iterations
mybin(iter=500,n=10, p=0.7)
```

### Aiv)
>
  1000 iteration
  
```{r}
# 1000 iterations
mybin(iter=1000,n=10, p=0.7)
``` 
  
### Av) and B
>
  10000 iteration
  
```{r}
# 10000 iterations
# 
mybin(iter=10000,n=10, p=0.7)
```

## C
```{r}
# creating binomial distribution using dbinom
obj = dbinom(1:10,size = 10,prob = 0.7)
round(obj,4)
```

#TASK 3

## A

```{r}
# sample of 5 marbles from a set of 12 white (1's) and 8 black (0's) marbles
M = c(rep(1,12),rep(0,8));
M
# Without replacement
sample(M, size = 5,replace = FALSE)
```

## B
```{r}
# With replacement
sample(c(1,0), size = 5, prob = c(12/20,8/20), replace = TRUE)
```

## C and D
```{r}
# call myhyper function
source('lab5-hypergeometric.R')
myhyper(iter =100, n = 5, N = 20, r = 12) # for 100 iterations
myhyper(iter =200, n = 5, N = 20, r = 12) # for 200 iterations
myhyper(iter =500, n = 5, N = 20, r = 12) # for 500 iterations
myhyper(iter =1000, n = 5, N = 20, r = 12) # for 1000 iterations
myhyper(iter =10000, n = 5, N = 20, r = 12) # for 10000 iterations

```

##E
```{r}
# Produce a hypergeometric distribution using dhyper
obj = dhyper(x = 0:5, m = 12, n = 8, k = 5)
round(obj,4)
```

#TASK 4

## A
>
  In lab5.r there is a function called mysample(). Look at the code and describe what you think it should do - record your response here.

>
  The code generates n samples of 1:10 with replacement, converts the sampled vector in to a factor by assigning levels 1:10. Next a table is created with each member of the sampled set placed under its corresponding level. Finally, a barplot of the sampled set is created. This process is repeated for iter iterations. The last line of the code delays its execution for a time time.
  
## B

```{r}
mysample=function(n, iter=10,time=0.5){
  for( i in 1:iter){
    #make a sample
    s=sample(1:10,n,replace=TRUE)
    # turn the sample into a factor
    sf=factor(s,levels=1:10)
    #make a barplot
    barplot(table(sf)/n,beside=TRUE,col=rainbow(10), 
            main=paste("Example sample()", " iteration ", i, " n= ", n,sep="") ,
            ylim=c(0,0.2)
    )
    
    #release the table
    Sys.sleep(time)
  }
}

mysample(n = 1000,iter = 30,time = 1)
```

##C

>
  Describe what you see here.

>  
  I noticed that with every iteration the height of the plots keep changing mostly about the relative frequency 0.10.
  

#TASK 5

```{r}
# ways of choosing 4 out of 8
choose(8,4)
# probability that y >4
# q = vector of quantile =4
ppois(4,lambda = 2, lower.tail = FALSE)
# probability Y=10 ~ NegBin (p=0.4,r =3)
mynbin=function(y,r,p){
  choose(y-1,r-1)*p^r*(1-p)^(y-r)
}
mynbin(10,3,0.4)

# Calculate probability of 8S in a binomial with 15 samples 
pbinom(8,15,0.4)
```

