---
title: "Lab2"
author: "JONATHAN IFEGUNNI"
date: "August 30, 2018"
output: 
  html_document:
    fig_caption: yes
    keep_md: yes
    number_sections: yes
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Task 1
```{r}
getwd()
```

# Task 2

```{r}
mpg.obj = read.csv("EPAGAS.csv")
head(mpg.obj)
```

# Task 3 

## 
```{r}
mpg = mpg.obj$MPG
z = (mpg-mean(mpg))/sd(mpg)
mean(z)
sd(z)
```

## 

>
  The values of mpg that are possible outliers are:
  
```{r}
mpg[abs(z)>=2 & abs(z)<=3] 
```

## 

> 
  The values of mpg that are defined as outliers are:
  
```{r}
mpg[abs(z)>3]
```
  
##

```{r}
dplot = ifelse(abs(z)>3, "Red",
        ifelse(abs(z)>=2 &abs(z)<=3,"Blue", "Black"))
library(lattice)
dotplot(z,col=dplot)
  
```

# Task 4

```{r}
boxplot(mpg,col="Green",data=mpg.obj, notch=TRUE, horizontal = TRUE)
title("Boxplot representation of MPG")

```

##

>
  Using Chebyshev's theorem the proportion of data within 2 standard deviations of the mean of the mpg data is 75%.

```{r}
k = 2; 
(1-1/(k)^2) 
```

##

>
  Using R the exact, proportion within 2 standard deviation of the mean is 96%.
  
```{r}
length(z[abs(z)<2])/length(z)
```

##
>
  No
  
##
>
  Using Empirical rule, approximately 95% of the data should be within 2 standard deviations of the mean.
  
##
>
  It corresponds
  
##
>
  Yes the Empirical rule is valid in this case because the data has a normal distribution.
